\documentclass[11pt,a4paper]{article}

\input{preamble}

\begin{document}

\def\contador{Lesson 6}
\input{title}

\section{Quotient Groups and Homomorphisms}

\subsection{Cosets and counting}

Let \((G,\cdot)\) be a group.


\begin{defi}
Let \(H\leq G\) and \(a,b\in G\).
Define \(\cong_l\) over \(G\) by 
\[a\cong_l b\iff a^{-1}b\in H.\]
Whenever \(a\cong_l b\) we say \(a\) is left congruent to \(b\) module \(H\).
Define \(\cong_r\) over \(G\) by 
\[a\cong_r b\iff ab^{-1}\in H.\]
Whenever \(a\cong_r b\) we say \(a\) is right congruent to \(b\) module \(H\).
\end{defi}

\begin{rem}
    If \(G\) is Abelian, \[a - b\in H\iff b-a\in H\]
    for any \(a,b\in G\).
    This is not true in general unless \(G\) is Abelian.
\end{rem}


\begin{teo}
    Let \(H\leq G\).
    \begin{enumerate}[label=(\roman*)]
        \item Both \(\cong_l\) and \(\cong_r\) are equivalence relations on $G$.
        
        \item The equivalence class of $a \in G$ under left congruence mod $H$ is the set
        \[
         aH = \{ah \mid h \in H\}
        \]
    \item The equivalence class of $a \in G$ under right  congruence mod $H$ is the set
    \[
    Ha = \{ha \mid h \in H\}  .
    \]


    \item For any \(a\in G\), \(\left|H a\right|=|H|=|a H|\).
    \end{enumerate}
\end{teo}

\begin{proof}
    Try it yourself (or classwork).
\end{proof}


We call \(aH\) the  \textit{left coset} of \(H\) by \(a\) in \(G\), and \(Ha\) the a \textit{right coset} of \(H\) by \(a\) in \(G\). 

\begin{rem}
    In additive notation (that is, when we are working with an Abelian group) we write \(a+H\) instead of \(aH\) and \(H+a\) instead of \(Ha\).
    In fact, there is no difference between left and right cosets in this case. (Why \(a+ H = H + a\) for any \(a\in G\)?)
\end{rem}
 



\begin{cor}\label{cor.hungerford.cor4.3}
Let \(H\leq G\).
\begin{enumerate}[label=(\roman*)]
    \item      \( G = \bigcup\limits_{a \in G} Ha = \bigcup\limits_{a \in G} aH \)
    \item For all \(a,b\in G\) distinct, \( aH \cap bH = \emptyset \) and \( Ha \cap Hb = \emptyset \).
    \item For all \(a,b\in G\), we have \( aH = bH \) if and only if \( a^{-1}b \in H \) (or \( b-a \in H \) in additive notation)  and \( Ha = Hb \) if and only if \( ab^{-1} \in H \) (or \( b-a  \in H \) in additive notation).
    \item If \( \mathcal{R} = \{ Ha \mid a \in G \} \) and \( \mathcal{L} = \{ aH \mid a \in G \} \) then \(|\mathcal{R}|=|\mathcal{L}|\).
\end{enumerate}
 
\end{cor}

A special name and notation have been adopted for the number of left (or right) cosets of a subgroup in a group.

\begin{defi}[Index]
    The index of a subgroup $H$ in $G$ is the number of distinct left cosets of $H$ in $G$. This number is denoted by $|G: H|$. 
\end{defi}

\begin{eje}
    Prove $|G: H|$ equals the number of distinct right cosets of $H$ in $G$.
    Thus it does not matter whether we count left or right cosets.
\end{eje}


\begin{defi}
    A \textbf{complete set of right representatives} of \(H\) is a subset \(S\) of \(G\) consisting of exactly one element from each right coset. In other words, 
    \(S\cap Ha \) is a singleton for every \(a\in G\).
\end{defi}

Define a \textit{complete set of left representatives} in the obvious way.
Note that such a set (either left or right) contains exactly one element of \(H\) since \(H = He\), where \(e\) is the identity of \(G\).
(What is the cardinality of a complete set of representatives?)
Further, if \(H=\langle e\rangle\), then \(Ha = \{a\} \) for any \(a\in G\), and \(|G:H| = |H|\), that is, there are as many left (or right) cosets as the number of elements of \(G\).

\begin{teo}
    If $K$, $H$, $G$ are groups  with $K<H<G$ then
\[
|G: K|=|G: H| \cdot|H: K|
\]
If any two of these indices are finite, so is the third.
\end{teo}

\begin{proof}
    Let \(\Lambda\) be a complete set of right representatives of \(H\) in \(G\).
    By Corollary \ref{cor.hungerford.cor4.3}, \[G=\bigcup_{a \in G} Ha =  \bigcup_{a \in \Lambda} Ha.\]
    Basically, we are joining all the equivalence classes given by right congruence modulo \(H\), and their union covers \(G\) (why?)
    Similarly, let \(\Omega\) be a complete set of right representatives of \(K\) in \(H\) and write \[H = \bigcup_{b \in \Omega } Kb  .\]
    Therefore, \[G = \bigcup_{a \in\Lambda} Ha =\bigcup_{a \in\Lambda} \left( \bigcup_{b \in \Omega } Kb \right)a = \bigcup_{(a,b) \in \Lambda\times \Omega} Kba\]
    (If you are not confortable with this, you should review the definition of a union over a multiple indexed family of sets. See \href{https://en.wikipedia.org/wiki/List_of_set_identities_and_relations#Definitions}{here} and \href{https://www.gornahoor.net/library/Halmos_NaiveSetTheory.pdf}{here} (page 35).)
    Let's now prove that the cosets \(Kba\) are mutually disjoint.
    Suppose \(Kba = Kb'a'\).
    Then \(ba = kb'a'\) for some \(k\in K\).
    Since \(b,b',k\in H\), have \(Ha = H ba = H kb'a' = H a'\), whence \(a=a'\) because we are working with complete sets of representatives.
    Thus \(b = kb'\).
    The same reasoning gives \(Kb = Kkb' = Kb'\) whence \(b=b'\).
    This proves the cosets \(Kba\) are pairwise disjoint.
    Finally, it follows that \(|G:K| = |\Lambda\times \Omega|\) by definition of index, and so  \[|G:K| = |\Lambda||\Omega| = |G:H| |H:K|,\]
    as desired.
    The last statement of the theorem is obvious.
\end{proof}


\begin{cor}[Lagrange's theorem]
    If \( H \leq G \), then \( |G| = [G : H] |H| \). 
In particular, if \( G \) is finite, the order of  any \( a \in G \) divides \( |G| \).
\end{cor}


\begin{proof}
    Apply the last theorem with $K=\langle e\rangle$ for the first statement. The second is a special case of the first with $H=\langle a\rangle$.
\end{proof}


If $G$ is a group and $H, K$ are subsets of $G$, we denote by $H K$ the set $\{a b \mid a \in H, b \in K\}$. Note that a right or left coset of a subgroup is a special case of this construction.

\begin{rem}
    Careful! If $H, K$ are subgroups, $H K$ may not be a subgroup.
\end{rem}

\begin{teo}
    Let \(H\) and \(K\) be finite subgroups of a group \(G\). Then \[|HK| =  \frac{|H||K|}{|H\cap K|}.\]
\end{teo}

\begin{proof}
    Note $C=H \cap K$ is a subgroup of $K$ of index $n=$ $|K| /|H \cap K|$ and $K$ is the disjoint union of right cosets $C k_1 \cup C k_2 \cup \ldots \cup C k_n$ for some $k_i \in K$. Since $H C=H$, this implies that $H K$ is the   union of the disjoint sets $H k_1 \cup H k_2 \cup$ $\cdots \cup H k_n$. Therefore, $|H K|=|H| \cdot n=|H||K| /|H \cap K|$.
\end{proof}


\noindent
\begin{minipage}[c]{0.5\textwidth}
\textbf{Proposition 1.9.} If \( H \) and \( K \) are subgroups of a group \( G \), then \( [H : H \cap K] \leq [G : K] \). If \( [G : K] \) is finite, then \( [H : H \cap K] = [G : K] \) if and only if \( G = HK \).
\end{minipage}%
\hfill
\begin{minipage}[c]{0.45\textwidth}
\[\begin{tikzcd}
    & G \arrow[ld, no head] \arrow[rd, no head] &                       \\
H \arrow[rd, no head] &                                           & K \arrow[ld, no head] \\
    & H\cap G                                   &                      
\end{tikzcd}\]
\end{minipage}

\begin{proof}
    Let $A$ be the set of all right cosets of $H \cap K$ in $H$ and $B$ the set of all right cosets of $K$ in $G$.
    The map $\varphi: A \rightarrow B$ given by $(H \cap K) h \mapsto K h$, with $h \in H$, is well defined since $(H \cap K) h^{\prime}=(H \cap K) h$ implies $h^{\prime} h^{-1} \in H \cap K \subset K$ and hence $K h^{\prime}=K h$.
    Complete the proof by following these simple steps:
    \begin{enumerate}[label=(\roman*)]
        \item Show that $\varphi$ is injective. Then $|H: H \cap K|=|A| \leq|B|$ $=|G: K|$.
        \item If $|G: K|$ is finite, then show that $|H: H \cap K|=|G: K|$ if and only if $\varphi$ is surjective.
        \item $\varphi$ is surjective if and only if $G=K H$.
    \end{enumerate}
    
    (Hint: note that for $h \in H$ and $k \in K$, we have $ K k h=K h$ since $(k h) h^{-1}=k \in K$.) Conclude.
\end{proof}


\begin{prop}
    Let \( H \) and \( K \) be subgroups of \( G \) with finite index of a group \( G \). Then \( |G : H \cap K| \) is finite and \(|G : H \cap K| \leq |G : H| |G : K| \). Furthermore, \(|G : H \cap K| = |G : H| |G : K| \) if \( G = HK \).
\end{prop}

\begin{proof}
    Classwork.
\end{proof}



\end{document}