\documentclass[
    11pt,a4paper,
    % answers
]{exam}

\input{preamble}

\begin{document}
\def\contador{Problem Set 5}
\input{title}

\begin{center}
    {\bfseries\Large
        Rings and their Properties}\\
    (Lesson 9)
\end{center}


\begin{questions}

\question
Let $A$ be a ring with unity. Show that if $u$ is invertible in $A$ then so is $-u$.


\question
 Prove that   $\Z_p$ is a integral domian iff $p$ is prime.

  
\question
The \emph{direct product} of the rings $A$ and $B$ is the cartesian product $A\times B$ endowed with the operations defined by
\begin{align*}
(x_1,y_1) + (x_2,y_2) &= (x_1+x_2, y_1+y_2),
\\
(x_1,y_1) \cdot (x_2,y_2) &= (x_1 x_2, y_1 y_2).
\end{align*}
\begin{enumerate}[label=(\roman*)]
\item Prove that $A\times B$ is a ring with these operations.
\item Prove that if $A$ and $B$ are Abelian, then so is its direct product.
\item Describe the divisors of zero of $A\times B$.
\item Describe the invertible elements of $A\times B$.
\item Assume that $A$ and $B$ are non trivial rings. Explain why $A\times B$ can not be an integral domain by providing a counterexample.
\end{enumerate}


% \question
% Let $(A_i)_{i\in I}$ be an  indexed family of rings. 
% \begin{enumerate}
%     \item Prove that $\prod_{i\in I} A_i$ is a ring under componentwise addition and multiplication. 
%     \item If $I=\Z^{+}\cup\adel{0},$ prove that $\coprod_{i\in I} A_i$  under componentwise addition and multiplication. 
%     \item Prove that $A\odel{x}$ is a ring. 
% \end{enumerate}


\question
Let $(A,+,\cdot)$ be a ring.
\begin{enumerate}[label=(\roman*)]
\item Prove that for any $a,b,c\in A$,
$$
a(b-c) = ab - ac,
\quad
(b-c)a = ba - ca.
$$
\item Assume that $a,b\in A$ are such that $ab=-ba$. Prove that 
$$
(a+b)^2 = (a-b)^2 = a^2 + b^2.
$$
\item Assume that $A$ is an integral domain. Prove that for all \(a,b\in A\), if \(a^2 = b^2\) then \(a=b\) or \(a=-b\).
\item If \(A\) is an integral domain, for any \(a\in A\), if  \(a=a^{-1}\), then \(x \in \{-1,1\}\). 
\item Prove that if $(A,+)$ is a cyclic group, then $A$ is a commutative ring.
\end{enumerate}



% \question
%  Let $A$ a non-void set equipped with internal operations $+$ and $\cdot$. Assume that $(A,+)$ is a group,
%  $(A,\cdot)$ is a semigroup, and that
% $$
% \forall a,b,c \in A:\quad
% a \cdot (b+c) = a\cdot b + a\cdot c
% \:\: \wedge \:\:
% (b+c)\cdot a = b\cdot a + c\cdot a;
% $$
% $$
% \exists 1\in A, \forall x\in A:\quad
% x\cdot 1 = 1\cdot x = 1.
% $$
% Prove that $A$ is a ring with unity.


\question
\label{ej:ejercicior2}
Let $(A,+,\cdot)$ be a nontrivial ring with unity and $a,b,c\in A$.
\begin{enumerate}[label=(\roman*)]
\item Prove that if $a$ is invertible, then 
$$
ab = ac \: \Rightarrow \: b=c,
$$
and that $a$ has only one multiplicative inverse.
%
\item Prove that if $a^2 = 0$, then $a+1$ and $a-1$ are invertible. 
\item Prove that if $a$ and $b$ are invertible, then $ab$ is invertible.
\item Prove that $(A^\times,\cdot)$ is a group.
\end{enumerate}


\question
Let $(F,+,\cdot)$ be a field with $|F|=m\in \mathbb{N}$. Prove that
\begin{equation}
\forall x\in F\setminus \{0\}:\quad
x^{m-1} = 1.
\end{equation}


\question
Let $A$ be a commutative ring and $a,b\in A$. Prove that if $ab$ is invertible, then $a$ and $b$ are both invertible.



\question
Let $(A,+,\cdot)$ be a nontrivial ring and $a,b,c\in A$.
\begin{enumerate}
\item Prove that if $a\notin \{-1,1\}$ and $a^2=1$, then $a+1$ and $a-1$ are zero divisors
\item Prove that if $ab$ is a divisor of zero, then either $a$ or $b$ is a zero divisor.
\end{enumerate}


\question
Prove that in a nontrivial commutative ring with unity, a zero divisor cannot be invertible. 




\question
Consider $A = (\Z, \oplus, \odot)$ where
$$
a\oplus b = a+b-1,
\quad
a\odot b = ab - (a+b) +2.
$$
\begin{enumerate}
\item Prove that $A$ is a commutative ring with unity. Indicate the zero element, the unit, and the negative of an arbitrary $a$.
\item Is $A$ an integral domain?
\end{enumerate}


\question
Consider $A = (\Q \times \Q, \oplus, \odot)$ where
$$
(a,b)\oplus (c,d) = (a+c,b+d),
\quad
(a,b)\odot (c,d) = (ac-bd,ad+bc).
$$
\begin{enumerate}[label=(\roman*)]
\item Prove that $A$ is a commutative ring with unity. Indicate the zero element, the unit element, and the negative of an arbitrary $a$.
\item Prove that $A$ is a field and indicate the multiplicative inverse of an arbitrary nonzero element.
\end{enumerate}


% \question
% Consider   $A = \Q(\sqrt{2})$.
% \begin{enumerate}[label=(\roman*)]
% \item Prove that $A$ is a commutative ring with unity. Indicate the zero element, the unit, and the negative of an arbitrary $a=x+y\sqrt{2}$.
% \item Prove that $A$ is a field.
% \end{enumerate}


\question
Explain why $\R^X$ is neither a field nor an integral domain.
Describe its  zero divisors.
Here \(X\) denotes an arbitary set.



%\question
%What kind of algebraic structure is $(\matriz[n](\real),+,\cdot)$? Explain why %$\matriz[n](\real)$ is not an integral domain.
%

\question
Let $\Omega \neq \varnothing$ be a set and consider $A = (\mathscr{P}(\Omega), \Delta, \cap)$.
\begin{enumerate}[label=(\roman*)]
\item Prove that $A$  is a commutative ring with unity.
\item Describe the zero divisors and the invertible elements of $A$.
\item Explain why $A$ is not an integral domain.
\item Give the tables for additiona and multiplication of $A$ for $\Omega = \{a,b,c\}$.
\end{enumerate}


% \question
% The set of \emph{quaternions}, $\quaternions$, can be seen as the elements of $\matriz[2](\complex)$ with the form
% \begin{equation}
% \label{eq:trium1}
% \alpha = 
% \begin{pmatrix}
% a+bi & c+di \cr
% -c+di & a-bi
% \end{pmatrix},\quad a,b,c,d \in \real.
% \end{equation}
% \begin{enumerate}
% \item Prove that $\quaternions$ endowed with the usual addition and multiplication of matrices is a non-commutative ring with unity.
% \item Prove that $\alpha$ as given in \eqref{eq:trium1}
% can be written, in \emph{standard notation}, as
% \begin{equation}
% \label{eq:fuimr1}
% \alpha = a \mathbf{1} + b \mathbf{i} + c\mathbf{j}
% + d \mathbf{k},
% \end{equation}
% where
% $$
% \mathbf{1} = \begin{pmatrix} 1 & 0 \cr 0 & 1    \end{pmatrix},\quad
% \mathbf{i} =
% \begin{pmatrix} i & 0 \cr 0 & -i    \end{pmatrix},\quad
% \mathbf{j}=
% \begin{pmatrix} 0 & 1 \cr -1 & 0    \end{pmatrix},\quad
% \mathbf{k} =
% \begin{pmatrix} 0 & i \cr i & 0    \end{pmatrix}.
% $$
% \item Prove that
% \begin{equation}
% \mathbf{i}^2 = \mathbf{j}^2 =
% \mathbf{k}^2 = \mathbf{1};
% \end{equation}
% \begin{equation}
% \mathbf{i} \mathbf{j} = - \mathbf{j} \mathbf{i} = \mathbf{k},\quad
% \mathbf{j} \mathbf{k} = - \mathbf{k} \mathbf{j} = \mathbf{i},\quad
% \mathbf{k} \mathbf{i} = - \mathbf{i} \mathbf{k} = \mathbf{j}.
% \end{equation}
% \item The \emph{conjugate} and \emph{norm} of the quaternion $\alpha = a \mathbf{1} + b \mathbf{i} + c\mathbf{j}
% + d \mathbf{k}$ are, respectively,
% \begin{eqnarray}
% \overline{\alpha} = a \mathbf{1} - b \mathbf{i} - c\mathbf{j}
% - d \mathbf{k},
% \\
% \|\alpha\| = \sqrt{a^2 + b^2 + c^2 + d^2}.
% \end{eqnarray}
% Prove that 
% $$
% \alpha \overline{\alpha} = \overline{\alpha} \alpha = \|\alpha\|^2 \mathbf{1}.
% $$
% \item Prove that $\quaternions$ is a \emph{skew field} i.e., it's a (not necessarily commutative) ring with unity in which every nonzero element has a multiplication inverse.
% \end{enumerate}


\question
Let $G$ be an additive Abelian group. An endomorphism on $G$ is a homomorphism from $G$ into $G$. Prove that $\mathrm{End}(G)$, the set of endomorphisms on $G$ becomes a ring with unity when it's endowed with addition and the composition product.


\question
Let $(A,+,\cdot)$ be a ring. An element $a\in A$ is said to be \emph{nilpotent} if
$$
\exists n\in \mathbb{N}:\quad a^n = 0.
$$
\begin{enumerate}[label=(\roman*)]
\item Prove that if $A$ has a unit element and $a\in A$ is nilpotent, then both $a+1$ and $a-1$ are invertible.
\item Prove that if $A$ is commutative and $a\in A$ is nilpotent, then $xa$ is nilpotent, for all $x\in A$.
\item Prove that if $A$ is commutative and $a,b\in A$  are nilpotent, then $a+b$ is nilpotent.
\end{enumerate}

\question
Show that a ring that is cyclic under addition is commutative. 


\question
Let $n$ be an integer greater than 1 . In a ring in which $x^n=x$ for all $x$, show that $a b=0$ implies $b a=0$.

\question
Let $a$ belong to a ring $R$. Let $S=\{x \in R \mid a x=0\}$. Show that $S$ is a subring of $R$.

\question
Suppose that there is a positive even integer $n$ such that $a^n=a$ for all elements $a$ of some ring. Show that $-a=a$ for all $a$ in the ring.


\question
Let $R$ be a ring. The center of $R$ is the set $\{x \in R \mid a x=x a$ for all $a$ in $R\}$. Prove that the center of a ring is a subring.


\question
Suppose that $R$ is a ring and that $a^2=a$ for all $a$ in $R$. Show that $R$ is commutative.


\question 
Show that $4 x^2+6 x+3$ is a unit in $Z_8[x]$.


\question
Let $R$ be a commutative ring with more than one element. Prove that if for every nonzero element $a$ of $R$ we have $aR = R$, then $R$ has a unity and every nonzero element has an inverse.

\question
Suppose that \(R\) is a ring with no zero divisors and that \(R\) contains a nonzero element \(x\) such that \(x^2 = x\).
Show that \(x\) is the unity of \(R\).

\question
Let \(R\) be a ring.
Prove that \(a^2 - b^2 = (a+b)(a-b)\) for all \(a,b\in R\) if and only if \(R\) is commutative.
\end{questions}

\end{document}