\documentclass[
    11pt,a4paper,
    % answers
]{exam}

\input{preamble}

\begin{document}
\def\contador{Problem Set 6}
\input{title}

\begin{center}
    {\bfseries\Large
        EDs, PIDs, UFDs, and Polynomial  Rings}\\
    (Lessons 12, 13 and  14)
\end{center}

\begin{questions}

\question
Let $R$ be an integral domain. Prove that if two elements $d$ and $d'$ of $R$ generate the same principal ideal, i.e., $\langle d \rangle = \langle d' \rangle$, then $d' = ud$ for some $u \in R^\times$. Conclude that if $d$ and $d'$ are both greatest common divisors of $a$ and $b$, then $d' = ud$ for some $u \in R^\times$.

\question
Consider the ring
\begin{align*}
R = \mathbb{Z}\left[\frac{1 + \sqrt{-d}}{2}\right] &= \left\{m + n\frac{1 + \sqrt{-d}}{2} \mid m, n \in \mathbb{Z} \right\}\\
&= \left\{\frac{a + b\sqrt{-d}}{2} \mid a, b \in \mathbb{Z}, a \equiv b \pmod{2} \right\}
\end{align*} 
where $d$ is square-free and $d \equiv 3 \pmod{4}$. Let $N(r) = r\bar{r}$, where $\bar{r}$ is the conjugate of $r$.
\begin{enumerate}[label=(\roman*)]
    \item Suppose that for any complex number $z$, there exists $\xi \in R$ such that $N(z - \xi) < 1$. Show that $N$ is a Euclidean function on $R$. Conclude that $R$ is a Euclidean domain.
    \item Show that $R = \mathbb{Z}\left[\frac{1 + \sqrt{-3}}{2}\right]$ is a Euclidean domain.
\end{enumerate}

\question Prove that every nonzero prime ideal in a Principal Ideal Domain
is a maximal ideal.

\question Prove that any two nonzero elements of a P.I.D. have a least common
multiple.

\question Let \(R\) be an integral domain and suppose that every prime ideal in \(R\)
is principal. Prove that every ideal of \(R\) is principal. i.e., \(R\) is a P.I.D.

\question
Show that the ideal $I = \langle 3, x^3 - x^2 + 2x - 1 \rangle$ in $\mathbb{Z}[x]$ is not principal.

\question
Find the greatest common divisor (GCD) of the polynomials $g = x^3 + x^2 + x - 3$ and $f = x^4 - x^3 + 3x^2 + x - 4$ in $\mathbb{Q}[x]$ using the Euclidean algorithm.

\question
Let $R = \mathbb{Z}[\sqrt{-5}]$.
\begin{enumerate}[label=(\roman*)]
    \item Prove that $2$, $\sqrt{-5}$, and $1 + \sqrt{-5}$ are irreducible in $R$.
    \item Prove that $R$ is not a unique factorization domain (UFD).
    \item Provide an explicit ideal in $R$ that is not principal.
\end{enumerate}

\question
Show that the polynomial
\[
(x - 1)(x - 2) \cdots (x - n) - 1
\]
is irreducible over $\mathbb{Z}$ for all $n \geq 1$.


\question
Show that the polynomial 
\[
(x - 1)(x - 2)(x - n) + 11
\]
is irreducible over $\mathbb{Z}$ for all $n \geq 1$, $n \neq 4$.

\question
Let $A[[x]]$ denote the set of sequences $(a_i)_{i \in \mathbb{Z}^+}$ of elements of $A$, without any restriction. Define the addition and multiplication of two elements $(a_i)_{i \in \mathbb{Z}^+}$ and $(b_i)_{i \in \mathbb{Z}^+}$ of $A[[x]]$ as follows:
\[
(a_i)_{i \in \mathbb{Z}^+} + (b_i)_{i \in \mathbb{Z}^+} = (a_i + b_i)_{i \in \mathbb{Z}^+}
\]
and
\[
(a_i)_{i \in \mathbb{Z}^+}(b_i)_{i \in \mathbb{Z}^+} = (c_k)_{k \in \mathbb{Z}^+},
\]
where $c_k = \sum_{i+j=k} a_i b_j$. Show that these operations make $A[[x]]$ a commutative ring and that $A[x]$ is a subring of $A[[x]]$. The ring $A[[x]]$ is called the \emph{ring of formal power series} in $x$: this name comes from the fact that, if we set $x = (0, 1, 0, \dots) \in A[[x]]$, then every element $(a_i)_{i \in \mathbb{Z}^+}$ of $A[[x]]$ can be formally written as:
\[
\sum_{i=0}^\infty a_i x^i.
\]

\question
Prove that $x^3 + nx + 2$ is irreducible in $\mathbb{Z}[x]$ for all integers $n \neq 1, -3, -5$.

\question
Determine whether the following polynomials are irreducible in the rings indicated. For those that are irreducible, determine their factorization into irreducibles.
\begin{enumerate}[label=(\roman*)]
    \item $x^3 + x + 1$ in $\mathbb{Z}_3[x]$.
    \item $x^5 + 1$ in $\mathbb{Z}_5[x]$.
    \item $x^4 + 10x^2 + 1$ in $\mathbb{Z}[x]$.
\end{enumerate}

\question
Apply Eisenstein's criterion to establish the irreducibility of each of the following polynomials over $\mathbb{Q}$:
\begin{multicols}{2}
    \begin{enumerate}[label=(\roman*)]
        \item $x^3 + 6x^2 + 2x + 2$,
        \item $x^5 - 2x^3 + 10$,
        \item $x^4 + 6$,
        \item $x^7 - 31$,
        \item $2x^4 - 27x^3 + 6x^2 - 9x + 6$,
        \item $x^3 + 5x^2 + 25x + 5$.
    \end{enumerate}
\end{multicols}


\question
Let $A$ be a ring. Prove that
\[
\frac{A[x,y]}{\langle x - y \rangle} \cong A[x] \cong A[y].
\]

\end{questions}

\end{document}